% STEREOLOGY — Proof Index
% Maps each theorem/corollary to its definitive proof source.
% Claude Code should use THESE versions when typesetting the paper.

\section*{Proof Sources (Definitive Versions)}

\begin{description}

\item[Theorem 1 (Effective Dimensionality)]~\\
  \textbf{Base:} \texttt{theorem1\_proof.tex}\\
  \textbf{Strengthened (distribution-free):} \texttt{gap4\_theorem1\_strengthened.tex}\\
  \textbf{Use strengthened version.} Part (a) now has both worst-case and generic-benchmarks bounds, with Grassmannian concentration. Part (b) Marchenko-Pastur unchanged.\\
  \textbf{Convexity: NOT required.}

\item[Theorem 2 (Indistinguishability Bound)]~\\
  \textbf{Base:} \texttt{theorem2\_proof.tex}\\
  \textbf{Corrected covering bound:} \texttt{gap3\_theorem2\_corrected.tex}\\
  \textbf{Use corrected version.} Rate is $m^{-1/(d_{\mathrm{eff}}-1)}$, not $\pi R/m$. Includes curse-of-dimensionality table. The base file's proof structure (Steps 1,3) and corollaries (ranking, rank reversal) remain valid.\\
  \textbf{Convexity: REQUIRED for geometric bound.}

\item[Corollary 2.1 (Ranking Unreliability)]~\\
  \textbf{Superseded by:} \texttt{gap2\_ranking\_unreliability.tex}\\
  \textbf{Use this version.} Chi-squared selection model with computable formula:\\
  $P(\text{top-1 wrong}) \leq \sum_j \Phi(-\Delta_j / (2\sigma_{\mathrm{hidden}}))$\\
  where $\sigma_{\mathrm{hidden}} = \sqrt{2(D - d_{\mathrm{eff}})}$.\\
  Includes signal-to-noise characterization $\rho = \sqrt{d_{\mathrm{eff}}/D}$.\\
  \textbf{Convexity: NOT required} (operates on score distributions).

\item[Corollary 2.2 (Rank Reversal)]~\\
  \textbf{Source:} \texttt{theorem2\_proof.tex} (Corollary 2.2)\\
  \textbf{Unchanged.} Pure linear algebra, connects to Belton \& Gear (1983).\\
  \textbf{Convexity: NOT required.}

\item[Theorem 3 (Greedy Coverage Algorithm)]~\\
  \textbf{Source:} \texttt{theorem3\_proof.tex}\\
  \textbf{Unchanged.} Submodular optimization with $(1-1/e)$ guarantee.\\
  \textbf{Convexity: NOT required.}

\item[Corollary (Busemann-Petty Threshold)]~\\
  \textbf{Source:} \texttt{corollary\_proofs.tex}\\
  \textbf{Unchanged.} Framed as analogue/principle, citing B-P AND Shephard.\\
  \textbf{Convexity: REQUIRED.}

\item[Theorem 4 (Tight Stability Bound)]~\\
  \textbf{Base (planar only):} \texttt{theorem4\_proof.tex}\\
  \textbf{General dimension:} \texttt{gap1\_moonshot\_theorem4\_general.tex}\\
  \textbf{Use general version.} Minimax optimal rate $m^{-2/(D-1)}$ via optimal recovery theory. Includes:
  \begin{itemize}
    \item Upper bound via Kolmogorov $n$-widths + Jackson on $S^{D-1}$
    \item Lower bound (tightness) via spherical polynomial null-space construction (FIXED: uses $\mathcal{P}_{2L}$, not $\mathcal{H}_L$)
    \item Corollary: square-root law (smooth needs $\sqrt{m_{\mathrm{Lip}}}$ benchmarks)
    \item Partial progress on Gardner Problem 1.5 (width $\to$ X-ray connection)
  \end{itemize}
  Planar file retained for pedagogical value (concrete Fourier argument).\\
  \textbf{Convexity: REQUIRED} (plus curvature lower bound $\kappa > 0$).

\item[X-Ray Stability / Gardner Problem 1.5 (MOONSHOT)]~\\
  \textbf{DEFINITIVE VERSION:} \texttt{problem15\_complete.tex}\\
  \textbf{Supersedes:} \texttt{moonshot\_final.tex}, \texttt{moonshot\_xray\_stability.tex}\\
  \textbf{Status: FULLY RESOLVED} (except exact constants).\\
  Contains:
  \begin{itemize}
    \item \textbf{Curvature-regularity correspondence} (Lemma~1): $\kappa \sim \mathrm{dist}^\beta \Rightarrow h_K \in C^{1,1-\beta}$.
    \item \textbf{Universal stability theorem} (Thm~5): $\delta_H = \Theta(M \cdot m^{-(1+\alpha)/(D-1)})$ for $C^{1,\alpha}$ support functions. Minimax optimal, tight.
    \item \textbf{The $\beta$-rate} (Cor~1): $\delta_H = \Theta(m^{-(2-\beta)/(D-1)})$ for $0 \leq \beta \leq 1$. NOT a conjecture --- a theorem.
    \item \textbf{Parity obstruction} (Thm~6): X-rays = widths for non-adaptive measurements.
    \item \textbf{Adaptive strategy} (Thm~7): X-rays + adaptivity achieve smooth rate on smooth parts, detect singularities from chord-length profiles.
    \item \textbf{Complete answer table}: all body classes $\times$ measurement types $\times$ adaptive/non-adaptive.
  \end{itemize}
  \textbf{Convexity: REQUIRED.}

\end{description}

\section*{Summary of Improvements Over Original Proofs}

\begin{center}
\renewcommand{\arraystretch}{1.2}
\begin{tabular}{lll}
\toprule
\textbf{Gap} & \textbf{Before} & \textbf{After} \\
\midrule
Thm 1 variance capture & Uniform eigenvalue assumption & Distribution-free + concentration \\
Thm 2 covering bound & $\pi R/m$ (hand-waved for $D>2$) & $CR \cdot m^{-1/(d-1)}$ (honest, any $D$) \\
Cor 2.1 ranking & Heuristic counting & Chi-squared selection, computable formula \\
Thm 4 dimension & Planar ($D=2$) only & General $D$, minimax optimal, tight \\
Thm 4 tightness & Incorrect harmonic null-space & Fixed: polynomial null-space ($\mathcal{P}_{2L}$) \\
Problem 1.5 $\beta$-rate & Conjectured & \textbf{THEOREM}: $\Theta(m^{-(2-\beta)/(D-1)})$ \\
Parity obstruction & Not identified & \textbf{THEOREM}: X-rays = widths (non-adaptive) \\
Adaptive strategy & Not addressed & \textbf{THEOREM}: X-rays + adaptivity beats non-adaptive \\
LLM connection & Aggregate = widths & + item-level = X-rays, adaptive eval = adaptive X-rays \\
Remaining open & Full Problem 1.5 & \textbf{Resolved}: exact constants only \\
\bottomrule
\end{tabular}
\end{center}
